---
title: "Part 3"
author: "Lauren Kennedy"
date: '2023-01-15'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Real data

```{r}
library(tidyverse)
library(brms)
library(lmeresampler)

data("jsp728")

df <- jsp728 %>%
  mutate(student_id = 1:n())%>%
  select(gender, class, school, normAge11, normAge8, student_id)%>%
  pivot_longer(cols = c(normAge11, normAge8),
               names_to = "time")

df_t1 <- df %>%
  filter(time == "normAge8")
```


Fit a multilevel model with a random effect for school and a fixed effect for class and gender

```{r}
ex3_model1 <- brm(value ~ gender + class + (1|school), data = df_t1, backend = "cmdstanr", cores = 4)

summary(ex3_model1)
```


Obtain fixed and random parameter estimates:

```{r}
fixef(ex3_model1)
ranef(ex3_model1)
```

Fit a fixed effects model for comparison

```{r}
ex3_model2 <- brm(value ~ gender + class + school, data = df_t1, backend = "cmdstanr", cores = 4)

fixef(ex3_model2)
```


Compare the median predicted estimate for each school in each model.

```{r}
ex3_pred1 <- data.frame(t(posterior_predict(ex3_model1)), school = df_t1$school, model = "random") 

ex3_pred2 <- data.frame(t(posterior_predict(ex3_model2)), school = df_t1$school, model = "fixed") 

combined_data <- rbind(ex3_pred1,ex3_pred2)

combined_data %>%
  pivot_longer(-c('school','model'),
               names_to = "posterior_id",
               values_to = "posterior_preds") %>%
  group_by(school,model)%>%
  summarise(y_med = median(posterior_preds), y_min = quantile(posterior_preds,.1), y_max = quantile(posterior_preds,.9))%>%
  ggplot(., aes(x = school,y=y_med, ymin = y_min, ymax = y_max, colour = model))+
  geom_point(position = position_dodge(width = .5)) +
  geom_errorbar(position = position_dodge(width = .5),width = 0)+
  theme_bw()+
  ggthemes::scale_color_colorblind()+
  coord_flip()
```


Fit students nested within schools


```{r}
ex3_model2 <- brm(value ~ gender + class + (1|school/student_id), data = df, backend = "cmdstanr", cores = 4)
```
# Exercises:

- Practice understanding the priors in these models using a prior predictive check
- Explore the suitability of the model using a posterior predictive check
- Set a very tight prior for the variance of the varying effect: do the estimates change for each school?
- Explore the data: do the number of observations in each school impact the amount of pooling?